Eco-evolutionary dynamics of finite populations from first principles

### 2023-03-24 ind_files/04_field_eqns.tex
Minor rephrasing

    \frac{\partial \phi}{\partial t}(x,t) = \mathcal{A}^{-}(x|\phi) + \frac{1}{\sqrt{K}}\mathcal{A}^{+}(x|\phi)\dot{W}(x,t)

@@

\section{The infinite population limit}
Taking $K \to \infty$ in equation \eqref{functional_langevin} then yields a PDE:
\begin{equation}
\label{deterministic_traj}
\frac{\partial \psi}{\partial t}(x,t) = \mathcal{A}^{-}\left(x|\psi\right) = b_K(x|\psi)- d_K(x|\psi)
\end{equation}
where I have used a different symbol $\psi$ simply to highlight that $\psi(x,t)$ as the solution to equation \eqref{deterministic_traj} is a deterministic function, whereas $\phi(x,t)$ as defined in equation \eqref{functional_langevin} is really a stochastic process $\{\phi^{(t)}\}_{t\geq0}$. Equation \eqref{deterministic_traj} simply says that in the absence of stochasticity, the change in the density of individuals with trait values $x$ is given by the difference between the birth and death rates of these individuals in the population. Models of this form are precisely the `PDE models' discussed in studies of Adaptive Diversification~\citep{doebeli_adaptive_2011}. Equation \eqref{deterministic_traj} is also the basic equation of `oligomorphic dynamics'~\citep{sasaki_oligomorphic_2011, lion_multimorph_2022} if one assumes the population is composed of a small number of `morphs', \emph{i.e.} $\psi(x,t) = \sum\limits_{k=1}^{S} n_{k}\psi_k(x,t)$, where $\psi_k(x,t)$ is the phenotypic distribution of the $k$th morph (often assumed a normal distribution with narrow variance) and $S$ is the number of distinct morphs in the population. Models of the form \eqref{deterministic_traj} are also used to study intraspecific trait variation in community ecology~\citep{nordbotten_dynamics_2020}. A prominent recent example is~\posscite{wickman_theoretical_2022} `trait space equations' in their framework for eco-evolutionary community dynamics.

>>>>

    \frac{\partial \phi}{\partial t}(x,t) = \mathcal{A}^{-}(x|\phi) + \sqrt{\frac{\mathcal{A}^{+}(x|\phi)}{K}}\dot{W}(x,t)

@@

\section{The infinite population limit}
Taking $K \to \infty$ in equation \eqref{functional_langevin} yields a PDE:
\begin{equation}
\label{deterministic_traj}
\frac{\partial \psi}{\partial t}(x,t) = \mathcal{A}^{-}\left(x|\psi\right) = b_K(x|\psi)- d_K(x|\psi)
\end{equation}
where I have used a different symbol $\psi$ simply to highlight that $\psi(x,t)$ as the solution to equation \eqref{deterministic_traj} is a deterministic function, whereas $\phi(x,t)$ as defined in equation \eqref{functional_langevin} is really a stochastic process $\{\phi^{(t)}\}_{t\geq0}$ in which each $\phi^{(t)}$ is a finite measure. Equation \eqref{deterministic_traj} simply says that in the absence of stochasticity, the change in the density of individuals with trait values $x$ is given by the difference between the birth and death rates of these individuals in the population. Models of this form are precisely the `PDE models' discussed in studies of Adaptive Diversification~\citep{doebeli_adaptive_2011}. Equation \eqref{deterministic_traj} is also the basic equation of `oligomorphic dynamics'~\citep{sasaki_oligomorphic_2011, lion_multimorph_2022} if one assumes the population is composed of a small number of `morphs', \emph{i.e.} $\psi(x,t) = \sum\limits_{k=1}^{S} n_{k}\psi_k(x,t)$, where $\psi_k(x,t)$ is the phenotypic distribution of the $k$th morph (often assumed a normal distribution with narrow variance) and $S$ is the number of distinct morphs in the population. Models of the form \eqref{deterministic_traj} are also used to study intraspecific trait variation in community ecology~\citep{nordbotten_dynamics_2020}. A prominent recent example is~\posscite{wickman_theoretical_2022} `trait space equations' in their framework for eco-evolutionary community dynamics.
